\documentclass[a4paper,9pt]{scrreprt}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsbsy}
\usepackage{graphicx}
\usepackage{vmargin}
\usepackage{array}
\usepackage{bm}
\usepackage{verbatim}
\usepackage[english]{babel}
\usepackage{color}
\usepackage{cite}
\usepackage{url}
\usepackage{fancyhdr}

\pagestyle{empty}

% Latex compact
\usepackage{mathptmx}
\linespread{0.9}
\addtolength{\topmargin}{-1cm}
\addtolength{\topskip}{-0.8cm}
\addtolength{\textheight}{1cm}
\parindent=0em
\addtolength{\parskip}{-1mm}
\renewcommand\floatpagefraction{.9}
\renewcommand\topfraction{.9}
\renewcommand\bottomfraction{.9}
\renewcommand\textfraction{.1}   
\setcounter{totalnumber}{50}
\setcounter{topnumber}{50}
\setcounter{bottomnumber}{50}

\begin{document}

\section*{{\huge Benjamin Menetrier}}
$  $\\
$  $\\
\large
\begin{tabular}{ll}
Birth date: & 05/14/1987 \\
Mail: &  Lieu-dit Rouvel\\
  & 12270 Najac, France \\
Tel.: & (+33) 7 77 04 32 19 \\
E-mail: & benjamin.menetrier@posteo.net \\
Website: & benjaminmenetrier.free.fr \\
GitHub: & github.com/benjaminmenetrier
\end{tabular}
$  $\\
\begin{center}
{\vrule depth 0pt height 0.4pt width 8cm}
\end{center}

\section*{Career}
\begin{itemize}
\item \textbf{October 2018 - present}: Research scientist at IRIT\footnote{Institut de Recherche en Informatique de Toulouse}, Toulouse, France. Subcontracting full time for the JCSDA\footnote{Joint Center for Satellite Data Assimilation}, Boulder, Colorado.\vspace{-0.1cm}
\item \textbf{April 2016 - September 2018}: Research scientist at CNRM\footnote{Centre National de Recherches Météorologiques}, Toulouse, France.\vspace{-0.1cm}
\item \textbf{October 2015 - March 2016}: Research engineer at CERFACS\footnote{Centre Européen de Recherche et de Formation Avancée en Calcul Scientifique}, Toulouse, France.\vspace{-0.1cm}
\item \textbf{Septembre 2014 - September 2015}: Project Scientist I at NCAR\footnote{National Center for Atmospheric Research}, Boulder, Colorado.
\end{itemize}

\section*{Education}
\begin{itemize}
\item \textbf{2011-2014}: PhD on the subject: "Use of an ensemble data assimilation to model flow-dependent background error covariances at convective scale", directed by Thibaut Montmerle, Yann Michel and Loïk Berre (CNRM/Météo-France), defended on July 3, 2014. PhD award ``Léopold Escande'' 2014 of the INPT\footnote{Institut National Polytechnique de Toulouse}.\vspace{-0.1cm}
\item \textbf{2011}: ENS\footnote{{\'E}cole Normale Supérieure, a prestigious institution of higher education providing specialized training to students who will become professors and researchers in their field} (Paris) graduation, geosciences specialty.\vspace{-0.1cm}
\item \textbf{2010}: Master OACOS\footnote{Ocean, Atmosphere, Climate and Remote Sensing}, in the Atmosphere, Ocean and Climate Dynamics specialty, obtained with highest distinction.\vspace{-0.1cm}
\item \textbf{2008}: Licence\footnote{Equivalent to Bachelor's degree} in Applied Physics to Earth Sciences, obtained with highest distinction.\vspace{-0.1cm}
\item \textbf{2007}: Admission to the ENS, Paris, France.\vspace{-0.1cm}
\item \textbf{2005-2007}: Classes Préparatoires\footnote{Classes Préparatoires are a two-year intensive course preparing for the competitive entrance examinations to French Grandes Écoles. Courses taken include scientific subjects such as Mathematics, Physics, and Chemistry but also Humanities and Languages.} in Dijon, France.\vspace{-0.1cm}
\item \textbf{2005}: Baccalauréat Scientifique \footnote{Equivalent to A-levels in Mathematics, Biology and Physics} obtained with highest distinction.
\end{itemize}

\section*{Other skills}
\begin{itemize}
\item Language: fluent English (over 22 months in English-speaking countries), French mother tongue.\vspace{-0.1cm}
\item Computer science: familiar with Windows, good knowledge of Linux operating systems.\vspace{-0.1cm}
\item Programming: extensive knowledge of FORTRAN, good working knowledge of C++, shell scripting and Python, word-processing with Latex.\vspace{-0.1cm}
\item Significant experience of development in many operational data assimilation systems, for the atmosphere or the ocean: ARPEGE-IFS\footnote{Action de Recherche Petite {\'E}chelle Grande {\'E}chelle-Integrated Forecasting System}, GSI\footnote{Gridpoint Statistical Interpolation}, JEDI\footnote{Joint Effort for Data assimilation Integration}, NEMOVAR, \footnote{Object-Oriented Prediction System} and WRF-DA\footnote{Weather Research and Forecasting-Data Assimilation}.
\end{itemize}

\section*{Teaching}
\begin{itemize}
\item \textbf{2019 - 2021}: lectures and practicals about the BUMP\footnote{Background error covariance on Unstructured Mesh Package} software for the ``JEDI academies'' providing training for the JEDI system users.\vspace{-0.1cm}
\item \textbf{2016 - 2021}: lectures about hybrid ensemble-variational methods for data assimilation at CERFACS, Toulouse.\vspace{-0.1cm}
\item \textbf{2018}: internship supervisor of two engineer and Master students (Toulouse).\vspace{-0.1cm}
\item \textbf{2016 - 2018}: supervisor of Engineering in Enterprise Projects at ISAE\footnote{Institut Supérieur de l'Aéronautique et de l'Espace}, Toulouse.\vspace{-0.1cm}
\item \textbf{2011-2014}: lectures at the ENM\footnote{{\'E}cole Nationale de la Météorologie}, Toulouse: oceanography, waves and swell, for technician and engineer students.\vspace{-0.1cm}
\item \textbf{2013}: co-supervisor of practical work at Paul Sabatier University, Toulouse: experimental and computational fluid dynamics.
\end{itemize}

\section*{Internships and scientific visits}
\begin{itemize}
\item \textbf{October 2018 - July 2019}: bi-annually visit at the JCSDA (Boulder, Colorado).
\item \textbf{August 2017 and 2018}: "B matrix bootcamp" for the development of the BUMP software, organized by the JCSDA (NCAR, Boulder, Colorado).\vspace{-0.1cm}
\item \textbf{May - August 2015}: Development of the EVIL\footnote{Ensemble-Variational Integrated Localized} algorithm in the operational GSI system, with Thomas Auligné (JCSDA, College Park, Maryland).\vspace{-0.1cm}
\item \textbf{September-October 2013}: Objective filtering of sampled background error covariances, with Thomas Auligné (NCAR, Boulder, Colorado).\vspace{-0.1cm}
\item \textbf{October 2010-June 2011}: Use of an ensemble data assimilation to model flow-dependent background error covariances at convective scale, with Thibaut Montmerle (CNRM/Météo-France, Toulouse, France).\vspace{-0.1cm}
\item \textbf{March-June 2010}: Use of heterogeneous forecast error covariances for the fog analysis in the fine scale weather prediction model AROME, with Thibaut Montmerle (CNRM/Météo-France, Toulouse, France).\vspace{-0.1cm}
\item \textbf{March-August 2009}: Data assimilation on an axisymmetric hurricane model with an ensemble Kalman filter, with Chris Snyder (NCAR, Boulder, Colorado)\vspace{-0.1cm}
\item \textbf{September 2008}: Study of the development of radiative fog based on the ParisFog campaign, with Thomas Dubos (LMD\footnote{Laboratoire de Météorologie Dynamique}-X, Palaiseau, France).\vspace{-0.1cm}
\item \textbf{June 2008}: Numerical study of 2D compressible turbulence, with Marie Farge (LMD-ENS, Paris, France).\vspace{-0.1cm}
\item \textbf{September 2008}: Numerical experimentation about data assimilation with variationnal methods, with Olivier Talagrand (LMD-ENS, Paris, France).
\end{itemize}

\section*{Publications}

\subsection*{Peer-reviewed}

\begin{itemize}
\item Marquet, P., P. Martinet, J.-F. Mahfouf, A. L. Barbu, and \textbf{B. Ménétrier}, 2021: Towards the use of conservative thermodynamic variables in data assimilation: preliminary results using ground-based microwave radiometer measurements. \textit{Atmospheric Measurement Techniques Discussions}, 1-20.
\item Weaver AT., M. Chrust, \textbf{B. Ménétrier}, and A. Piacentini, 2020: An evaluation of methods for normalizing diffusion-based covariance operators in variational data assimilation. \textit{Quarterly Journal of the Royal Meteorological Society}, 1-32.
\item Martinet, P., D. Cimini, F. Burnet, \textbf{Ménétrier, B.}, Y. Michel, and V. Unger, 2020: Improvement of numerical weather prediction model analysis during fog conditions through the assimilation of ground-based microwave radiometer observations: a 1D-Var study. \textit{Atmospheric Measurement Techniques}, \textbf{13}, 6593-6611.
\item Montmerle, T. , Y. Michel, E. Arbogast,  \textbf{B. Ménétrier}, and P. Brousseau, 2018: A 3D Ensemble Variational Data Assimilation Scheme for the limited area AROME model: formulation and preliminary results. \textit{Quarterly Journal of the Royal Meteorological Society}, \textbf{144(716)}, 2196-2215.\vspace{-0.1cm}
\item Laloyaux, P. , S. Frolov, \textbf{B. Ménétrier}, and M. Bonavita, 2018: Implicit and explicit cross‐correlations in coupled data assimilation. \textit{Quarterly Journal of the Royal Meteorological Society}, \textbf{144(715)}, 1851-1863.\vspace{-0.1cm}
\item Berre, L., E. Arbogast, \textbf{B. Ménétrier}, and G. Desroziers, 2017: Change of variable applied to mass and wind fields for covariance localisation. \textit{WMO CAS/JSC WGNE Blue Book}, Edited by J. Côté.\vspace{-0.1cm}
\item Michel, Y., \textbf{B. Ménétrier} and T. Montmerle, 2016: Objective Filtering of the Local Correlation Tensor. \textit{Quarterly Journal of the Royal Meteorological Society}, \textbf{142(699)}, 2314-2323. \vspace{-0.1cm}
\item Auligné, T., \textbf{B. Ménétrier}, A.C. Lorenc and M. Buehner, 2016: Ensemble-Variational Integrated Localized Data Assimilation. \textit{Monthly Weather Review}, \textbf{144(10)}, 3677-3696.\vspace{-0.1cm}
\item Bouttier, F., L. Raynaud, O. Nuissier, and \textbf{B. Ménétrier}, 2015: Sensitivity of the AROME ensemble to initial and surface perturbations during HyMeX. \textit{Quarterly Journal of the Royal Meteorological Society}, \textbf{142(S1)}, 390-403.\vspace{-0.1cm}
\item \textbf{Ménétrier, B.} and T. Auligné, 2015: An overlooked issue of variational data assimilation. \textit{Monthly Weather Review}, \textbf{143(10)}, 3925-3930.\vspace{-0.1cm}
\item \textbf{Ménétrier, B.} and T. Auligné, 2015: Objective optimization of localized-hybridized ensemble-based covariances. \textit{Monthly Weather Review}, \textbf{143(10)}, 3931-3947.\vspace{-0.1cm}
\item \textbf{Ménétrier, B.}, T. Montmerle, Y. Michel, and L. Berre, 2015: Linear filtering of sample covariances for ensemble-based data assimilation. Part I: Optimality criteria and application to variances filtering and covariances localization. \textit{Monthly Weather Review}, \textbf{143(5)}, 1622-1643.\vspace{-0.1cm}
\item \textbf{Ménétrier, B.}, T. Montmerle, Y. Michel, and L. Berre, 2015: Linear filtering of sample covariances for ensemble-based data assimilation. Part II: Application to a convective-scale NWP model.  \textit{Monthly Weather Review}, \textbf{143(5)}, 1644-1664.\vspace{-0.1cm}
\item \textbf{Ménétrier, B.}, T. Montmerle, L. Berre, and Y. Michel, 2014: Estimation and diagnosis of heterogeneous flow-dependent background error covariances at convective scale using either large or small ensembles. \textit{Quarterly Journal of the Royal Meteorological Society}, \textbf{140(683)} , 2050-2061.\vspace{-0.1cm}
\item \textbf{Ménétrier, B.} and T. Montmerle, 2011: Heterogeneous background-error covariances for the analysis and forecast of fog events. \textit{Quarterly Journal of the Royal Meteorological Society}, \textbf{137(661)}, 2004-2013.\vspace{-0.1cm}
\item Montmerle, T.,  Y. Michel, and \textbf{B. Ménétrier}, 2011: Modelling of background error covariances for the analysis of clouds and precipitation. \textit{ECMWF-JCSDA Workshop on Assimilating Satellite Observations of Clouds and Precipitation into NWP Models}, 15-17 June 2010, 121-132.\vspace{-0.1cm}
\item Nguyen van Yen, R., B. Kadoch, V. Kumar, \textbf{B. Ménétrier}, M. Farge, K. Schneider, D. Douady, and L. Guez, 2011: Influence of waves on Lagrangian acceleration in two-dimensional turbulent flows. \textit{ESAIM: Proc}, \textbf{32}, 231-241.
\end{itemize}

\subsection*{Theoretical and technical documentation (available on my GitHub page)}
\begin{itemize}
\item Sample covariance filtering. DOI: 10.5281/zenodo.4054856
\item Multivariate localization. DOI: 10.5281/zenodo.4009125
\item NICAS documentation. DOI: 10.5281/zenodo.4058620
\item Vertical balance equations. DOI: 10.5281/zenodo.5728709
\end{itemize}

\section*{Symposiums, conference and  seminars}
Invited:\vspace{0.2cm}
\begin{itemize}
\item \textbf{Ménétrier, B.}, 2019: Localization for ensemble DA: objective diagnostic and efficient application. \textit{Workshop on Climate Prediction in the Atlantic-Arctic sector}, 5-7 June 2017, Bergen, Norway.\vspace{-0.1cm}
\item \textbf{Ménétrier, B.}, 2017: Normalized Convolution Interpolated from an Adaptive Subgrid. \textit{JCSDA B matrix bootcamp}, 31 July-18 August 2017, Boulder, Colorado.\vspace{-0.1cm}
\item \textbf{Ménétrier, B.}, 2017: Developments in ensemble variational data assimilation at Météo-France. \textit{ECMWF Seminar}, 25 April 2017, Reading, United Kingdom.\vspace{-0.1cm}
\item \textbf{Ménétrier, B.} and T. Auligné, 2015: Optimized localization and hybridization to filter ensemble-based covariances. \textit{NASA/GMAO seminar}, 21 July 2015, Greenbelt, Maryland.\vspace{-0.1cm}
\item \textbf{Ménétrier, B.}, T. Auligné, A. Lorenc, et M. Buehner, 2015: Un nouveau schéma d’assimilation variationnelle ensembliste: EVIL. \textit{Séminaire du CNRM}, 27 avril 2015, Toulouse, France.\vspace{-0.1cm}
\item \textbf{Ménétrier, B.} and T. Auligné, 2015: An EVIL look at data assimilation. \textit{International workshop on theoretical aspects of ensemble data assimilation for the Earth system}, 5-10 April 2015, Les Houches, France.\vspace{-0.1cm}
\item \textbf{Ménétrier, B.}, T. Montmerle, Y. Michel, and L. Berre, 2014: Linear filtering of sample covariances for ensemble data assimilation: theory and applications of optimality criteria. \textit{NCAR/MMM seminar}, 28 August 2014, Boulder, Colorado.
\end{itemize}
$  $\\
Speaker:\vspace{0.2cm}
\begin{itemize}
\item \textbf{Ménétrier, B.}, 2021: Recent developments of the BUMP software. \textit{JCSDA Technical Review Meeting and Science Workshop}, 7-11 June 2018, virtual.\vspace{-0.1cm}
\item \textbf{Ménétrier, B.}, G. Vernières, T. Sluka, 2021: Application of the BUMP library
in the SOCA system. \textit{Joint ECMWF/OceanPredict workshop on Advances in Ocean Data
Assimilation}, 17-20 May 2021, virtual.\vspace{-0.1cm}
\item \textbf{Ménétrier, B.}, 2020 : BUMP, a generic package for background error covariance modeling. \textit{Fourth Data Assimilation Task Team Meeting}, Toulouse, France.\vspace{-0.1cm}
\item \textbf{Ménétrier, B.}, E. Arbogast, L. Berre, and Y. Trémolet, 2018: The Normalized Interpolated Convolution on an Adaptive Subgrid (NICAS) method, a new implementation of localization for EnVar applications. \textit{Workshop on Sensitivity Analysis and Data Assimilation in Meteorology and Oceanography}, 1-6 July 2018, Aveiro, Portugal.\vspace{-0.1cm}
\item \textbf{Ménétrier, B.}, 2016: Optimisation conjointe de la localisation et de l’hybridation pour filtrer des covariances échantillonnées. \textit{Colloque National sur l'Assimilation de données}, 30 novembre-2 décembre 2016, Grenoble, France.\vspace{-0.1cm}
\item \textbf{Ménétrier, B.}, T. Montmerle, Y. Michel, L. Berre, T. Auligné et A. Weaver, 2016: Localisation et hybridation de covariances échantillonnées pour l'assimilation de données. \textit{Journées GMMC 2016}, 7-9 juin 2016, Toulon, France.\vspace{-0.1cm}
\item \textbf{Ménétrier, B.} and T. Auligné, 2015: Optimized localization and hybridization to filter ensemble-based covariances. \textit{Tenth workshop on sensitivity analysis and data assimilation in meteorology and oceanography}, 1-5 June 2015, Roanoke, West Virginia.\vspace{-0.1cm}
\item \textbf{Ménétrier, B.}, T. Montmerle, Y. Michel, and L. Berre, 2014: Linear filtering of sample covariances for ensemble data assimilation: theory and applications of optimality criteria. \textit{World weather open science conference}, 16-21 August 2014, Montréal, Canada.\vspace{-0.1cm}
\item \textbf{Ménétrier, B.}, T. Montmerle, L. Berre, et Y. Michel, 2012: Variances d'erreur d'ébauche dépendantes de la situation métérologique à méso-échelle. \textit{Colloque national sur l'assimilation de données}, 17-19 décembre 2012, Nice, France.
\end{itemize}
$  $\\
Poster:\vspace{0.2cm}
\begin{itemize}
\item \textbf{Ménétrier, B.}, 2016: Optimized localization to filter ensemble-based covariances, applied to a coupled system. \textit{International workshop on coupled data assimilation}, 18-21 October 2016, Toulouse, France.\vspace{-0.1cm}
\item \textbf{Ménétrier, B.} and T. Auligné, 2016: Optimized localization and hybridization to filter ensemble-based covariances. \textit{International Symposium on Data Assimilation}, 18-22 July 2016, Reading, United Kingdom.\vspace{-0.1cm}
\item \textbf{Ménétrier, B.}, T. Montmerle, L. Berre, and Y. Michel, 2013: Spatial filtering of small ensemble-based estimations of background error parameters at convective scale. \textit{Sixth WMO symposium on data assimilation}, 7-11 October 2013, College Park, Maryland.\vspace{-0.1cm}
\item \textbf{Ménétrier, B.}, T. Montmerle, L. Berre, and Y. Michel, 2012: Adaptative denoising of ensemble-based background error variance maps at convective scale. \textit{International conference on ensemble methods in geophysical sciences}, 12-16 November 2012, Toulouse, France.\vspace{-0.1cm}
\item \textbf{Ménétrier, B.}, T. Montmerle, L. Berre, and Y. Michel, 2011: Filtering of ensemble-based forecast error variance maps, a toy-models approach. \textit{Ninth international workshop on adjoint model applications in dynamic meteorology}, 10-14 October 2011, Cefalu, Italy.
\end{itemize}
$  $\\
Last update: \today.
\end{document}
